---
title: "ST2195 Coursework"
author: "Thon Hui Min"
date: "28/01/2023"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Setup

## Library

```{r}
# Import library
library (dplyr)
library (ggplot2)
library (skimr)
library (tidyr)
library(scales)
```

## Read CSV

```{r}
# Read Csv
flight_2000 <- read.csv("2000.csv", header = TRUE)
flight_2001 <- read.csv("2001.csv", header = TRUE)

airport <- read.csv("airports.csv")
carriers <- read.csv("carriers.csv")
plane_data <- read.csv("plane-data.csv", na.strings = "")
```

## Bind flights

```{r}
flights <- rbind(flight_2000, flight_2001)
```

## Data Summary

```{r}
summary(flights)
```

```{r}
summary(plane_data)
```

# Question 1

When is the best time of day, day of the week, and time of year to fly to minimise delays?

```{r}
#Introduce new variable total delay as delay
flights <- flights %>% 
  mutate(delay = ArrDelay + DepDelay) %>% 
  filter(delay >= 0) %>% 
  arrange(delay)

head(flights)
```

## Q1.1

Best Time of day

```{r}
#Time
best_time <- flights %>% 
  filter(delay > 0) %>% 
  group_by(CRSDepTime) %>% 
  summarise(avg_delay = mean(delay, na.rm = TRUE)) %>% 
  arrange(avg_delay)

best_time
```

```{r}
## Plot
best_time %>% 
 ggplot(aes(x = CRSDepTime, y = avg_delay, colour = CRSDepTime)) + geom_line() + labs(title = 'Average Flight Delay by Time', x = 'Departure Time', y = "Average delay (in minutes)")
```

```{r}
## Timing with lowest delay
lowest_time <- best_time %>% 
  slice(which.min(avg_delay)) %>% 
  pull(CRSDepTime)

cat("The time with the lowest average flight delay is", lowest_time)
```

Hence, the best time of day to fly is 05:19 for the least delay.

## Q1.2

Best Day of the week to fly

```{r}
#Day
best_day <- flights %>% 
  filter(delay > 0) %>% 
  group_by(DayOfWeek) %>% 
  summarise(avg_delay = mean(delay, na.rm = TRUE)) %>% 
  arrange(avg_delay)
best_day
```

```{r}
# Graph
best_day %>% 
  mutate(DayOfWeek = factor(DayOfWeek, levels = c(1, 2, 3, 4, 5, 6, 7),
  labels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>% 
  ggplot() + 
  geom_col(aes(x = DayOfWeek, y = avg_delay, fill = 'Red'), width = 0.5) + 
  labs(title = 'Average Flight Delay by Day', x = 'Day of Week', y = 'Average Delay (in minutes)') + 
  theme(legend.position = "none") + 
  scale_fill_manual(values = "#0369a0")
```

```{r}
# Day with lowest average delay
lowest_day <- best_day %>% 
  slice(which.min(avg_delay)) %>% 
  pull(DayOfWeek)

cat("The day of the week with the lowest average flight delay is", weekdays(as.Date(lowest_day, origin = "1970-01-04")), ".")
```

Hence, the best day of the week to fly is: Tuesday

## Q1.3

Best month of the year to fly

```{r}
# Month
best_month <- flights %>% 
  filter(delay >= 0) %>% 
  group_by(Month) %>% 
  summarise(avg_delay = mean(delay, na.rm = TRUE)) %>% 
  arrange(avg_delay)
```

```{r}
head(best_month)
```

```{r}
# Bar Graph
ggplot(best_month, aes(x = Month, y = avg_delay)) +
  geom_col(fill = "#0369a0", width = 0.5) +
  labs(title = "Average Flight Delay by Month", x = "Month", y = "Average Delay (in minutes)") +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  theme_minimal() 
```

```{r}
# Month with lowest delay
lowest_month <- best_month %>% 
  filter(avg_delay == min(avg_delay)) %>% 
  pull(Month)

cat("The month with the lowest average flight delay is", month.name[lowest_month], ".")
```

Hence the best month of the year to fly is October.

# Question 2

Do older planes suffer more delays?

```{r}
#Remove all missing values
plane_data <- na.omit(plane_data)
head(plane_data)
```

```{r}
#Getting the manufacturing year based on TailNum on flights data frame
year_planes <- flights %>% 
    left_join(plane_data, by = c("TailNum" = "tailnum")) %>% 
    group_by(year, TailNum) %>% 
    summarise(avg_delay = mean(delay)) %>% 
    filter(avg_delay > 0, year != "None" & year != "NA" & year != "0000") %>% 
    arrange(year)
```

```{r}
year_planes
```

```{r}
# Scatter plot
year_planes %>% 
  ggplot() + aes(x = year, y = avg_delay, color = factor(year)) + geom_point(lwd = 2) + scale_x_discrete(breaks = seq(1956, 2007, by = 1)) + labs(title = "Number of Delays by Manufacturing Years", x = "Manufacturing Year", y = "Average Delay (in minutes)") + theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 90))
```

# Question 3

How does the number of people flying between different locations change over time?

```{r}
# Question 3
num_flights <- flights %>% 
  filter(delay > 0, flights$DepTime <= 2359, flights$CRSDepTime <= 2359,    flights$ArrTime <= 2359, flights$CRSArrTime <= 2359, !is.na(Year), !is.na(Month)) %>%
  group_by(Year, Month) %>%
  count()

flightstats <- flights %>% 
  filter(delay > 0) %>% 
  group_by(Year) %>% 
  count(Year)

flightstats
```

```{r}
num_flights
```

```{r}
# Bar Graph
ggplot(num_flights, aes(x = Month, y = n, fill = factor(Year))) + 
  geom_bar(position = position_dodge(width = 0.9), stat = "identity") + 
  scale_x_continuous(breaks = seq(1, 12, by = 1), labels = month.abb) + 
  scale_y_continuous(labels = scales::comma) + 
  labs(title = "Flights over the years", x = "Month", y = "Count", fill = 'Year') + 
  scale_fill_manual(values = c("#e24a33", "#348abd")) +
  theme(legend.position = "bottom") +
  geom_bar(data = num_flights %>% filter(Year == 2008),
           aes(x = Month, y = n, fill = "2008"), 
           position = position_dodge(width = 0.4), stat = "identity") +
  geom_bar(data = num_flights %>% filter(Year == 2009),
           aes(x = Month, y = n, fill = "2009"), 
           position = position_dodge(width = 0.4), stat = "identity")
```

# Question 4

Can you detect cascading failures as delays in one airport create delays in others?

```{r}
#Correlation coefficient between 2 variables
cor(flights$DepDelay, flights$ArrDelay, method = 'pearson', use = 'complete.obs')
```

```{r}
# Define custom color palette
my_colors <- c("#0072B2", "#E74C3C")

# Create scatter plot with regression line
ggplot(flights, aes(x = DepDelay, y = ArrDelay, color = factor(Year))) +
  geom_point(alpha = 0.2) +
  stat_smooth(method = "lm", formula = y ~ x, se = FALSE, size = 1.2, color = my_colors[2]) +
  scale_color_manual(values = my_colors) +
  labs(title = "Arrival Delay vs. Departure Delay",
       x = "Departure Delay (minutes)",
       y = "Arrival Delay (minutes)",
       color = "Year") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))
```

# Question 5

Use the available variables to construct a model that predicts delays.

```{r}
#Skim the data frame for missing values
skim(flights)
```

```{r}
#Drop missing values
flights_cleaned <- flights %>% 
  drop_na(DepTime, ArrTime, AirTime, ArrDelay, DepDelay) %>% 
  select (Month, DayOfWeek, CRSDepTime, CRSArrTime, ActualElapsedTime, DepDelay, Distance, ArrDelay) %>% 
  filter(CRSDepTime <= 2359, CRSArrTime <= 2359)
```

## Import Library

```{r}
#Machine learning preparation
library(mlr3)
library(mlr3learners)
library(mlr3pipelines)

#Ridge regression
library(glmnet)
library(paradox)
```

## Set Up

```{r}
# Task
task <- TaskRegr$new(id='flights', backend=flights_cleaned, target='ArrDelay')
measure <- msr('regr.mse')
```

```{r}
# Learner
learner_lm <- lrn('regr.lm')
```

```{r}
#Set seed and conduct train test split to get MSE score of Linear regression Model
set.seed(1)
gr_lm <- po(learner_lm)
glrn_lm <- GraphLearner$new(gr_lm)
train_set <- sample(task$nrow, 0.7 * task$nrow) 
test_set <- setdiff(seq_len(task$nrow), train_set)
glrn_lm$train(task, row_ids = train_set)
lm_score <- glrn_lm$predict(task, row_ids = test_set)$score() 
```

## Ridge Regressopm

```{r}
# Ridge Regression
lnr_ridge <- lrn('regr.glmnet')
lnr_ridge$param_set$values <-list(alpha= 0, lambda = 0.3)
gr_ridge <- po(lnr_ridge)
glrn_ridge <- GraphLearner$new(gr_ridge)
glrn_ridge$train(task,row_ids = train_set)
ridge_score <- glrn_ridge$predict(task, row_ids = train_set)$score()
```

```{r}
#Skim cleaned flights data frame
skim(flights_cleaned)
```

## Lasso

```{r}
#Lasso
lnr_lasso <- lrn('regr.glmnet')
lnr_lasso$param_set$values <-list(alpha= 1, lambda = 0.3)
gr_lasso <- po(lnr_lasso)
glrn_lasso <- GraphLearner$new(gr_lasso)
glrn_lasso$train(task,row_ids = train_set)
lasso_score <- glrn_lasso$predict(task, row_ids = train_set)$score()
```

## Random Forest Model

```{r}
# Import Library
library(ranger)
library(mlr3tuning)
library(mlr3learners)
```

```{r}
lrn_rf <- lrn('regr.ranger')
lrn_rf$param_set$values <- list(min.node.size = 4)
gr_rf <- po(lrn_rf)
glrn_rf <- GraphLearner$new(gr_rf)

tune_ntrees <- ParamSet$new(list(ParamInt$new('regr.ranger.num.trees', lower = 10, upper = 20)))

# Create a tuner and terminator object
tuner <- TunerRandomSearch$new()
terminator <- trm("evals", n_evals = 10)

# Define resampling with 70% train and 30% test split
resampling <- rsmp("holdout", ratio = 0.7)

AT_rf <- AutoTuner$new(learner = glrn_rf, resampling = resampling, measure = measure, search_space = tune_ntrees, terminator = terminator, tuner = tuner)

# Train using AutoTuner
AT_rf$train(task)

# Evaluate the model
prediction <- AT_rf$predict(task)

# Calculate the score
rf_score <- prediction$score(measure)
```

```{r}
print(lm_score)
print(ridge_score)
print(lasso_score)
print(rf_score)
```


## Graph

```{r}
# Create a data frame containing the scores
model_scores <- data.frame(
  Model = c("Linear Regression", "Ridge Regression", "Lasso Regression", "Random Forest"),
  MSE = c(lm_score, ridge_score, lasso_score, rf_score)
)

# Load the ggplot2 library for plotting
library(ggplot2)

# Create a bar chart comparing the MSE scores of each model
ggplot(model_scores, aes(x = Model, y = MSE, fill = Model)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = comma(MSE)), vjust = -0.5) +
  scale_y_continuous(labels = comma, limits = c(0, max(model_scores$MSE) * 1.1)) +
  theme_minimal() +
  labs(title = "Comparison of Regression Algorithms", x = "Model", y = "Mean Squared Error") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, vjust = 0.5)) +
  scale_x_discrete(limits = model_scores$Model)

```
